### 2020-12-18 deprecated/jds-paper-todo.tex

\section{todo}

We might want to have two documents; one with "everything" -- the other with highlights referencing first

\begin{verbatim}

In conclusions/highlights we could include

a) Data sample

b) simple remarks on data such as correlations between demographics and 
   that case/fatality data is quite rough and this limits insights

c) magic fits illustrate that DL can get good results when input data smooth

d) basic DL with just case and death data

e) DL+Demographics with clear but modest improvement limited (probably) by data quality

f) technical issues in DL

    Comparison with other DL for time series

    Errors and why use sqrt(count)

    Alternative normalization by population

    Method to combine case/death with demographics

    Calculating future values and custom loss function

    Dropout

    Hyper-parameter search

g) Sensitivity and other discussions of demographics 
h) Futures

What is related work?

\end{verbatim}

\subsection{Geoffrey}

...